Titeln på avhandlingens språk

### 2022-11-03 background.tex
greedy

Take \( \{ t_0, t_1, \ldots, t_{n + 1} \} \) a series of discrete and sequential sampling intervals (following a greedy scheme) at which the \ac{WCA} captures the state of the physical system, as illustrated in \cref{fig:step}.
\( t_0 \) corresponds to the instant at which the instruction for step \( M \) is provided and the first sample is taken, and \( t_n \) to the instant at which the final sample (i.e. which captures the final state of step \( M \)) is taken. 
\( t_{n + 1} \) is then the instant at which the result for sample \( t_n \) is returned and the instruction for step \( M + 1 \) is provided.
If we define \( t_c \) as the point in time at which the user finishes performing the instruction for step \( M \), it must follow that (see also~\cite{olguinmunoz:impact2021})\footnote{\( U(a, b) \) represents the continuous uniform distribution in the open interval \( (a, b) \).}
\begin{align}\label{eq:tc}
    t_c &\thicksim U(t_{n - 1}, t_n)
\end{align}

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/correlation_neuro_exectime.png}
    \caption{%
        Correlation between neuroticism and the mean execution time of the last four steps in segments of \num{12} steps subject to the same \ac{TTF} in~\textcite{olguinmunoz:impact2021}.
        Pearson correlation coefficient \( \rho = 0.418 \), 2-tailed \( p < 0.05 \).
    }\label{fig:neuroexectimecorrel}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/mu_fits_exgaussian_slice0.png}
    \caption{%
        \( \mu \) parameter of \acs*{exGaussian} distributions fitted to execution times of the first four steps of segments of steps subject to the same \ac{TTF} in \textcite{olguinmunoz:impact2021}.
        Fits were performed using the \ac{MLE} method.
    }\label{fig:muexgaussian}
\end{figure}

>>>>

Take \( \{ t_0, t_1, \ldots, t_{n + 1} \} \) a series of discrete and sequential sampling intervals (following a greedy scheme) at which the \ac{WCA} captures the state of the physical system, as illustrated in \cref{fig:step}.
\( t_0 \) corresponds to the instant at which the instruction for step \( M \) is provided and the first sample is taken, and \( t_n \) to the instant at which the final sample (i.e. which captures the final state of step \( M \)) is taken. 
\( t_{n + 1} \) is then the instant at which the result for sample \( t_n \) is returned and the instruction for step \( M + 1 \) is provided.
If we define \( t_c \) as the point in time at which the user finishes performing the instruction for step \( M \), it must follow that (see also~\cite{olguinmunoz:impact2021})\footnote{\( U(a, b) \) represents the continuous uniform distribution in the open interval \( (a, b) \).}
% \todo[inline]{I think it's an assumption rather than a must. Specifically, it is not uniform if the general execution time distribution is Rayleigh/ExpGaussian -vishnu}
\begin{align}\label{eq:tc}
    t_c &\thicksim U(t_{n - 1}, t_n)
\end{align}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/correlation_neuro_exectime.png}
    \caption{%
        Correlation between neuroticism and the mean execution time of the last four steps in segments of \num{12} steps subject to the same \ac{TTF} in~\textcite{olguinmunoz:impact2021}.
        Pearson correlation coefficient \( \rho = 0.418 \), 2-tailed \( p < 0.05 \).
    }\label{fig:neuroexectimecorrel}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/mu_fits_exgaussian_slice0.png}
    \caption{%
        \( \mu \) parameter of \acs*{exGaussian} distributions fitted to execution times of the first four steps of segments of steps subject to the same \ac{TTF} in \textcite{olguinmunoz:impact2021}.
        Fits were performed using the \ac{MLE} method.
    }\label{fig:muexgaussian}
\end{figure}

### 2022-11-03 implications_optimization.tex
writing
writing intro
greedy

\begin{figure}
    \centering
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/lifetime_all_ttfs.png}
        \caption{%
            Evolution of mean application lifetimes as \acp{TTF} increase.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/lifetime_diff.png}
        \caption{%
            Percentage difference in mean application lifetimes with respect to the reference model at select \acp{TTF}.
            Error bars indicate the \SI{95}{\percent} \acp{CI}, calculated using a two-sided T-test.
        }
    \end{subfigure}
    \caption{\acl{TTF} versus application lifetime.}\label{fig:lifetimes}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/lifetime_testbed.png}
        \caption{%
            Mean application lifetimes per testbed configuration.
            Note that due to the low number of samples, means have been calculated using the geometric instead of arithmetic average. 
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using bootstrapping.
        }
    \end{subfigure}
    \par\bigskip
    \begin{subtable}{\columnwidth}
        \centering
        \begin{tabular}{lrrrr}
            \toprule
            \# clients & 1 & 4 & 7 & 10 \\
            \ac{RTT} & \SI{0.42}{\second} & \SI{1.12}{\second} & \SI{1.92}{\second} & \SI{2.68}{\second} \\
            \bottomrule
        \end{tabular}
        \caption{Mean measured \aclp{RTT} for each testbed configuration.}
    \end{subtable}
    \par\bigskip
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/lifetime_testbed_diff.png}
        \caption{%
            Percentage difference in mean application lifetimes with respect to the reference model.
            Confidence intervals have been omitted due to the low number of samples and the use of the geometric mean.
        }
    \end{subfigure}
    \caption{Application lifetimes in the realistic scenarios.}\label{fig:testbed_lifetimes}
\end{figure}

@@

@@

\( R_\text{max} \) and \( R_\text{min} \) correspond to the maximum and minimum sampling rates of the system, respectively.
\( R_\text{max} \) can directly be assumed to correspond to one over the mean \ac{RTT} of the system, whereas \( R_\text{min} \) needs to either be calculated according to the latency bounds of the system or specified manually.
\( \alpha \) corresponds to a scaling factor and \( t \) to the time of the current sampling instant with respect to the start of the step.
Finally, \( CDF \) corresponds to the \acl*{CDF} of a distribution describing the execution times for the current step; \citeauthor{Wang2019Towards} used a single static Gaussian distribution for all steps in their work.

In the following, we will show the effects of more accurate execution time estimation using our timing models on the performance of this adaptive scheme.
For this, we implement two variants of the adaptive scheme.
The first one corresponds to \citeauthor{Wang2019Towards}'s original design, using a Gaussian distribution fitted to all the execution times collected for~\cite{olguinmunoz:impact2021} for \ac{CDF} calculation.
The second variant instead uses an embedded internal execution time model to keep track of the \acp{TTF} of the system and adjust the distribution of execution times following the state of the timing model.

We proceed to set up an experiment where these sampling scheme variants are deployed on identical, simulated, \num{100}-step tasks.
The execution times for each step are generated by a realistic model parameterized with high neuroticism and without any distribution fitting on the data (this timing model is separate from the internal timing models inside the sampling schemes).
We set \( R_\text{min} \) to \SI{0.5}{\hertz}, \( \alpha \) to \num{3.0}, and the \ac{RTT} of the system to a constant \SI{150}{\milli\second}, and then run \num{90} repetitions of the task for the original design of the sampling scheme as well as for each possible configuration of it with a different parameterization of the timing model.

\begin{figure}
    \centering
    \begin{subtable}[]{\columnwidth}
        \centering
        \begin{tabular}{rrr}
            \toprule
            \( R_\text{min} \) & \( \alpha \) & \ac{RTT} \\
            \midrule
            \SI{0.5}{\hertz} & \num{3.0} & \SI{150}{\milli\second} \\
            \bottomrule
        \end{tabular}
        \caption{Parameters used for \textcite{Wang2019Towards}'s adaptive sampling scheme.}
    \end{subtable}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/sampling_junjue.png}
        \caption{%
            Mean number of samples per step over a \num{100}-step task, \num{90} independent repetitions per configuration.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=.9\textwidth]{figs/new_model/diff_sampling_junjue.png}
        \caption{%
            Percentage difference in mean number of samples per step.
            Error bars indicate the \SI{95}{\percent} \acp{CI}, calculated using a two-sided T-test.
        }
    \end{subfigure}
    \caption{Results of parameterizing \textcite{Wang2019Towards}'s adaptive sampling scheme using the realistic timing model.}\label{fig:samples}
\end{figure}

The results of this investigation are presented in \cref{fig:samples}, and they clearly show the potentially dramatic effects of more accurate execution time estimation when optimizing sampling rates.
When using the realistic model for execution time estimation, the sampling scheme achieve between \SI{22}{\percent} and \SI{27}{\percent} better average performance.
This is true even for configurations with low neuroticism which do not exactly match the parameterization of the external timing model driving the execution times for the steps, showing thus the importance of even less-accurate approximations.

@@

    =\Big(F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)\Big)\\

@@

>>>>

\begin{figure}
    \centering
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/lifetime_all_ttfs.png}
        \caption{%
            Evolution of mean application lifetimes as \acp{TTF} increase.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/lifetime_diff.png}
        \caption{%
            Percentage difference in mean application lifetimes with respect to the reference model at select \acp{TTF}.
            Error bars indicate the \SI{95}{\percent} \acp{CI}, calculated using a two-sample T-test.
        }
    \end{subfigure}
    \caption{\acl{TTF} versus application lifetime.}\label{fig:lifetimes}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/lifetime_testbed.png}
        \caption{%
            Mean application lifetimes per testbed configuration.
            Note that due to the low number of samples, means have been calculated using the geometric instead of arithmetic average. 
            Error bars indicate \SI{95}{\percent} \acp{CI}, calculated using bootstrapping.
        }
    \end{subfigure}
    \par\bigskip
    \begin{subtable}{\columnwidth}
        \centering
        \begin{tabular}{lrrrr}
            \toprule
            \# clients & 1 & 4 & 7 & 10 \\
            \ac{RTT} & \SI{0.42}{\second} & \SI{1.12}{\second} & \SI{1.92}{\second} & \SI{2.68}{\second} \\
            \bottomrule
        \end{tabular}
        \caption{Mean measured \aclp{RTT} for each testbed configuration.}
    \end{subtable}
    \par\bigskip
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/lifetime_testbed_diff.png}
        \caption{%
            Percentage difference in mean application lifetimes with respect to the reference model.
            Confidence intervals have been omitted due to the low number of samples and the use of the geometric mean.
        }
    \end{subfigure}
    \caption{Application lifetimes in the realistic scenarios.}\label{fig:testbed_lifetimes}
\end{figure}

@@

\todo[inline]{%
    Ergo we need a second order model. Everything that comes afterwards is only with second order model. Play with RTT.
}

@@

\( R_\text{max} \) and \( R_\text{min} \) correspond to the maximum and minimum sampling rates of the system, respectively.
\( R_\text{max} \) can directly be assumed to correspond to \( 1 / \text{\ac{RTT}}_\mu \), where \( \text{\ac{RTT}}_\mu \) corresponds to the mean \ac{RTT} of the system.
\( R_\text{min} \) needs to either be calculated according to the latency bounds of the system or specified manually, and \( \alpha \) corresponds to a scaling factor and \( t \) to the time of the current sampling instant with respect to the start of the step.
Finally, \( CDF \) corresponds to the \ac{CDF} of a distribution describing the execution times for the current step; \citeauthor{Wang2019Towards} used a single static Gaussian distribution for all steps in their work.

In the following, we will show the effects of more accurate execution time estimation using our timing models on the performance of this adaptive scheme.
For this, we implement two variants of the adaptive scheme.
The first one corresponds a reference implementation using \citeauthor{Wang2019Towards}'s original design, using a Gaussian distribution fitted to all the execution times collected for~\cite{olguinmunoz:impact2021} for \ac{CDF} calculation.
This variant the same \ac{CDF} for every step.
The second variant instead uses an embedded internal execution time model to keep track of the \acp{TTF} of the system and adjust the distribution of execution times following the state of the timing model.
In other words, in this variant the \ac{CDF} used for instantaneous calculation of the sampling rate \( R \) changes over time according to the state of the execution time model.
For timing model configurations which do not fit a distribution to execution time data, the \ac{CDF} corresponds to the \ac{ECDF} of the selected execution times.

We proceed to set up an experiment where these sampling scheme variants are deployed on identical, simulated, \num{100}-step tasks.
Also included for reference in this experiment is greedy (also known as \emph{zero-wait}) sampling, where samples are taken as fast as possible; a new sample is taken as soon as the result for the previous one has been received.
This is the default sampling strategy used in most existing \ac{WCA} prototypes.
The execution times for each step are generated by a realistic model parameterized with high neuroticism and without any distribution fitting on the data (this timing model is separate from the internal timing models inside the sampling schemes).
We set \( R_\text{min} \) to \SI{0.5}{\hertz}, \( \alpha \) to \num{3.0}, and the \ac{RTT} of the system to a constant \SI{150}{\milli\second}, and then run \num{90} repetitions of the task for the original design of the sampling scheme as well as for each possible configuration of it with a different parameterization of the timing model.

\begin{figure}
    \centering
    \begin{subtable}[]{\columnwidth}
        \centering
        \begin{tabular}{rrr}
            \toprule
            \( R_\text{min} \) & \( \alpha \) & \ac{RTT} \\
            \midrule
            \SI{0.5}{\hertz} & \num{3.0} & \SI{150}{\milli\second} \\
            \bottomrule
        \end{tabular}
        \caption{Parameters used for \textcite{Wang2019Towards}'s adaptive sampling scheme.}
    \end{subtable}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_junjue.png}
        \caption{%
            Mean number of samples per step over a \num{100}-step task, \num{90} independent repetitions per configuration.
            Error bars indicate \SI{95}{\percent} \acp{CI}.
        }
    \end{subfigure}
    \begin{subfigure}[]{\columnwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/diff_sampling_junjue.png}
        \caption{%
            Percentage difference in mean number of samples per step.
            Error bars indicate the \SI{95}{\percent} \acp{CI}, calculated using a two-sample T-test.
            \todo[inline]{%
                Add greedy.
                Develop optimal sampling scheme?
                Explain grouping better?
            }
        }
    \end{subfigure}
    \caption{%
        Results of parameterizing \textcite{Wang2019Towards}'s adaptive sampling scheme using the realistic timing model.
        \emph{ED2} indicates a sampling scheme using the \edgedroid{} model for \ac{CDF} estimation.
        \emph{HN} and \emph{LN} indicate high or low neuroticism, respectively, and \emph{ExG fit} indicates a timing model which fits an \ac{exGaussian} distribution to the execution time data for each step.
    }\label{fig:samples}
\end{figure}

The results of this investigation are presented in \cref{fig:samples}
They clearly show the potentially dramatic effects of more accurate execution time estimation when optimizing sampling rates.
When using the realistic model for execution time estimation, the sampling schemes achieve between \SI{22}{\percent} and \textasciitilde\SI{25}{\percent} better average performance.
This is true even for configurations with low neuroticism which do not exactly match the parameterization of the external timing model driving the execution times for the steps, showing thus the importance of even less-accurate approximations.

@@

    =\Big(F_\mathcal{T}(t_n)-\\\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)\Big)\\

@@

\subsubsection{Evaluating the impact of the realistic timing model on the energy-optimal sampling scheme}

\todo[inline]{
    TODO.

    Plot energy consumption over RTT, mean energy per step over a task.
    Ground truth empirical high.
    For greedy, ideal, junjues sampling, vishnus model with gt as estimator.
}

### 2022-11-03 model.tex
greedy

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/ttf_to_exectime.png}
    \caption{%
        Effects of feeding three different \acp{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}) into the model on the generated execution times.
        Higher \acp{TTF} directly lead to higher execution times.
        Error bars indicate the \SI{95}{\percent} \ac{CI}.
    }\label{fig:ttf_to_exectime}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.9\columnwidth]{figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \ac{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \ac{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bars indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:exectimeduration}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/ttf_to_exectime.png}
    \caption{%
        Effects of feeding three different \acp{TTF} (\emph{low}, \SI{0}{\second}, \emph{medium}, \SI{2.5}{\second}, or \emph{high}, \SI{5}{\second}) into the model on the generated execution times.
        Higher \acp{TTF} directly lead to higher execution times.
        Error bars indicate the \SI{95}{\percent} \ac{CI}.
    }\label{fig:ttf_to_exectime}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/exectime_over_steps.png}
    \caption{%
    Effects of prolonged exposure to constant levels of system impairment on the model.
    At low (\SI{0}{\second}) \ac{TTF}, the models speed-up over time; conversely at high (\SI{5.0}{\second}) \ac{TTF}, the models either present no change or drastically increase their generated execution times, depending on the level of neuroticism.
    Error bars indicate \SI{95}{\percent} \ac{CI}.
    }\label{fig:exectimeduration}
\end{figure}
